---
title: "Coursera Practical Machine Learning Project"
author: "Matthew Pang"
date: "December 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background and Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Data Processing

### Load library

```{r load library}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(knitr)
library(repmis)
```


```{r Set seed}
set.seed(520)
```
### Import Data
We download the data from the given URLs below:-

```{r import data}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

```
The training dataset consists of `r dim(training)[1]` of observations and `r dim(training)[2]` of variables.
The testing dataset consists of `r dim(testing)[1]` of observations and `r dim(testing)[2]` of variables.

### Data Cleaning
We remove the columns which have missing values

```{r Data cleaning}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```
## Data Spliting
```{r DataSpliting}
inTrain <- createDataPartition(training$classe, p=0.7, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]

trainData <- myTraining[, -c(1:7)]
testData <- myTesting[, -c(1:7)]

```

## Prediction Algorithms
We use classification trees and random forests to predict the outcome.

### Classification Trees
We use 5-fold cross to implement the algorithm

```{r classification trees}
control <- trainControl(method = "cv", number = 5)
fit_rpart <- train(classe ~ ., data = trainData, method = "rpart", 
                   trControl = control)
print(fit_rpart, digits = 4)
```

Plot the classification trees

```{r classification trees plot}
fancyRpartPlot(fit_rpart$finalModel)
```

Predict the outcome using testData
```{r Predict outcome using testData}
predict_rpart <- predict(fit_rpart, testData)

(conf_rpart <- confusionMatrix(testData$classe, predict_rpart))
```

Accuracy of the outcome
```{r Accuracy}
(accuracy_rpart <- conf_rpart$overall[1])
```

### Random forests
The classification tress does not perform well, hence we predict by using random forests method.

```{r random forests}
fit_rf <- train(classe ~ ., data = trainData, method = "rf", 
                trControl = control)
print(fit_rf, digits = 4)
```

Predict the trainData set.
```{r}
predict_rf <- predict(fit_rf, trainData)
(conf_rf <- confusionMatrix(trainData$classe, predict_rf))
```

Accuracy of random forests.
```{r accuracy of random forests}
(accuracy_rf <- conf_rf$overall[1])
```

Predict the testData set
```{r predict testData}
(predict(fit_rf, testing))
```

## Conclusion
This project proved that random forests perform better than classification trees in this scenario.